---
title: "BCB 520 Assignment 3"
format: html
editor: visual
---

**Assignment: Data Visualization using Reddit Hyperlink Network**

For this assignment, I’m creating a data visualization using the **Reddit Hyperlink Network dataset** from **Stanford’s SNAP** ([source](https://snap.stanford.edu/data/soc-RedditHyperlinks.html)). This dataset tracks how different subreddits are connected through hyperlinks—basically, when one subreddit links to another in a post.

I picked the second dataset to analyze subreddit interactions based on title mentions rather than mentions in in posts for simplicity. I'm using **Social Network Analysis (SNA)** to explore how communities interact and connect on Reddit. The goal is to visualize these relationships and see patterns in how subreddits reference each other!

```{r}
library(readr)

file_path <- "soc-redditHyperlinks-title.tsv"

# using read_tsv to read the data without needing to decompress it
reddit_data <- read_tsv(file_path, comment = "#")

head(reddit_data)

```

To understand this, we’ll **turn the data into a network (graph)** where:\
- **Each subreddit is a dot (node).**\
- **A line (edge) connects two dots if one subreddit links to another.**

To see how different communities are connected and which ones are the most linked or influential.

-   **SOURCE_SUBREDDIT** is where the link starts (the subreddit that posted the link).\
-   **TARGET_SUBREDDIT** is where the link goes (the subreddit being linked to).

```{r}
library(igraph)

edges <- reddit_data[, c("SOURCE_SUBREDDIT", "TARGET_SUBREDDIT")]

reddit_graph <- graph_from_data_frame(edges, directed = TRUE)

summary(reddit_graph)
```

This tells us: - The network consists of **54,075 subreddits**. - There are **571,927 hyperlinks** connecting them. - The graph is **directed**, meaning links have a source and a destination.

# Finding the most refrenced subreddits (in-degree centrality)

```{r}
top_influential <- sort(degree(reddit_graph, mode = "in"), decreasing = TRUE)[1:10]
cat("Top 10 Most Linked Subreddits (In-Degree Centrality):\n")
print(top_influential)

```

# Finding the most active subreddits (out-degree centrailty)

```{r}
top_outbound <- sort(degree(reddit_graph, mode = "out"), decreasing = TRUE)[1:10]
cat("Top 10 Subreddits That Link the Most (Out-Degree Centrality):\n")
print(top_outbound)

```

# Creating a network of the most refrenced subreddits

```{r}
# Get the top 50 most linked subreddits by in-degree
top_50 <- names(sort(degree(reddit_graph, mode = "in"), decreasing = TRUE))[1:50]

# Create a subgraph with only those subreddits
subreddit_subgraph <- induced_subgraph(reddit_graph, vids = top_50)

# Convert to tidygraph for visualization
library(ggraph)
library(tidygraph)

set.seed(42)
ggraph(as_tbl_graph(subreddit_subgraph), layout = "kk") + 
  geom_edge_link(aes(alpha = 0.5), color = "gray") +
  geom_node_point(size = 5, aes(color = degree(subreddit_subgraph, mode = "in"))) +
  geom_node_text(aes(label = name), repel = TRUE, size = 4) +
  theme_void() +
  ggtitle("Top 50 Most Linked Subreddits")

```
