---
title: "Anscombe's Quartet Visualization"
format:
  html:
    toc: true
    toc-depth: 2
    html-math-method: katex
    code-copy: true
---
For this assignment, I’m doing a visualization with Anscombe’s Quartet—a very interesting dataset that shows how important it is to actually *see* our data, not just look at the stats. It is named after Frank Anscombe, t he one who designed the dataset in 1973 to teach this lesson. 

The dataset consists of four groups of $x$ and $y$ values that share almost identical summary statistics (mean, variance, and correlation) but look so different when plotted. I’ll create two visualizations, one straightforward and coded by hand, and another with a little help from AI to make it more playful and polished.

# 1. Loading the R packages needed and tidying up the dataset,
```{r, message=FALSE, warning=FALSE}
library("tidyverse", "broom", "knitr")

data(anscombe)

anscombe_tidy <- anscombe %>%
  pivot_longer(cols = everything(),
               names_to = c(".value", "set"),
               names_pattern = "(.)(.)") %>%
  mutate(set = as.numeric(set))

view(anscombe_tidy)
```
# 2. Getting the summary stats, and we can see they are almost identical!
```{r}
summary_stats <- anscombe_tidy %>%
  group_by(set) %>%
  summarize(
    mean_x = mean(x),
    mean_y = mean(y),
    var_x = var(x),
    var_y = var(y),
    correlation = cor(x, y)
  )

print(summary_stats)
```
# 3. A plot to bring the data to life,
```{r, message=FALSE, warning=FALSE}
ggplot(anscombe_tidy, aes(x = x, y = y)) +
  geom_point(color = "blue", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  facet_wrap(~set, ncol = 2) +
  labs(
    title = "Anscombe's Quartet",
    x = "X Values",
    y = "Y Values"
  ) +
  theme_minimal()

```
## A nicer cuter one,

```{r, warning=FALSE, message=FALSE}
ggplot(anscombe_tidy, aes(x = x, y = y)) +
  geom_point(aes(color = "Data Points"), size = 4, alpha = 0.8, shape = 21, fill = "#80CBC4") +
  geom_smooth(aes(color = "Trend Line"), method = "lm", se = FALSE, size = 1.2, linetype = "dashed") +
  facet_wrap(~set, ncol = 2) +
  scale_color_manual(
    values = c("Data Points" = "#4DB6AC", "Trend Line" = "#E57373"),
    name = NULL
  ) +
  labs(
    title = "Anscombe's Quartet",
    subtitle = "Identical statistical summaries, visually distinct datasets. Like whaaat! :o",
    x = "X Values",
    y = "Y Values",
    caption = "Data Source: Anscombe (1973)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5, color = "#37474F"),
    plot.subtitle = element_text(size = 14, hjust = 0.5, color = "#546E7A"),
    strip.text = element_text(face = "bold", color = "#FF7043"),
    axis.title = element_text(face = "bold", color = "#37474F"),
    panel.grid.major = element_line(color = "#CFD8DC"),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "#F9FBE7", color = NA),
    legend.position = "top",
    legend.justification = "center",
    legend.box.spacing = unit(0.2, "cm")
  )

```