---
title: "Hockey Analytics"
subtitle: "Assignment 5: NHL Player Award Ballots"
format:
  html:
    toc: true
    toc-depth: 2
    html-math-method: katex
    code-copy: true
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
  error: false
---

In this assignment, I will create data driven ballots for 6 NHL awards, ranking five players per category. These awards such as the Hart Memorial Trophy for MVP, the Vezina Trophy for top goaltender, and the Calder Memorial Trophy for best rookie are determined by subjective voting, but my approach will rely on relevant performance metrics.

# Reading the data and basic joins and filtering

```{r}
library(tidyverse)
library(kableExtra)
library(pillar)


base_url <- "https://raw.githubusercontent.com/ProfessorPolymorphic/BCB520/main/posts/A5-TabularData/"

rookie_goalies      <- read_csv(paste0(base_url, "RookieGoalies.csv"))
rookie_skaters_ind  <- read_csv(paste0(base_url, "RookieSkaterIndividual.csv"))
rookie_skaters_onice<- read_csv(paste0(base_url, "RookieSkaterOnIce.csv"))
skaters_ind         <- read_csv(paste0(base_url, "SkaterIndividualstats.csv"))
skaters_onice       <- read_csv(paste0(base_url, "SkaterOnIcestats.csv"))

```

```{r}
all_skaters <- skaters_onice %>%
  inner_join(skaters_ind, by = c("Player", "Team", "Position", "GP"))

rookie_skaters <- rookie_skaters_onice %>%
  inner_join(rookie_skaters_ind, by = c("Player", "Team", "Position", "GP"))

```

# Hart Memorial Trophy

This is for player most valuable to the team. The criteria I will use is total points per game in at least 30 games played.

```{r}
mvp_candidates <- all_skaters %>%
  filter(GP >= 30) %>%
  mutate(PPG = `Total Points` / GP) %>%
  arrange(desc(PPG)) %>%
  select(Player, Team, Position, GP, `Total Points`, PPG)

#ballot
mvp_top5 <- head(mvp_candidates, 5)

```

# Vezina Trophy

This is for the top goaltender. I will use saves and goals against as metrics.

```{r}
vezina_candidates <- rookie_goalies %>%
  filter(GP >= 30) %>%
  mutate(SavePercentage = Saves / (Saves + `Goals Against`)) %>%
  arrange(desc(SavePercentage)) %>% 
  select(Player, Team, GP, Saves, `Goals Against`, SavePercentage)

#ballot
vezina_top5 <- head(vezina_candidates, 5)

```

# Calder Memorial Trophy

This is for rookie of the year, the most proficient player in his first year of competition.

```{r}
calder_candidates <- rookie_skaters %>%
  filter(GP >= 30) %>%
  mutate(PPG = `Total Points` / GP) %>%
  arrange(desc(PPG)) %>%
  select(Player, Team, Position, GP, `Total Points`, PPG)

#ballot
calder_top5 <- head(calder_candidates, 5)

```

# James Norris Memorial Trophy

This is awarded to the defenseman with the greatest all around ability. I will filter by Position == "D" and use a hypothetical Impact metric, which will be the total of points and assists relative to time on ice.

```{r}
norris_candidates <- all_skaters %>%
  filter(Position == "D", GP >= 30) %>%
  mutate(Impact = (`Total Points` + `Total Assists`) / TOI.x) %>% 
  arrange(desc(Impact)) %>%
  select(Player, Team, GP, `Total Points`, `Total Assists`, Impact)

#ballot
norris_top5 <- head(norris_candidates, 5)
```

# Frank J. Selke Trophy

This is for the forwards player who best excels in the defensive aspects of the game. I will be using on ice shot for percentage.

```{r}
selke_candidates <- all_skaters %>%
  filter(Position %in% c("C", "R"), GP >= 30) %>%
  mutate(DefensiveRating = `SF%`) %>%
  arrange(desc(DefensiveRating)) %>%
  select(Player, Team, GP, `Total Points`, `SF%`, DefensiveRating)

selke_top5 <- head(selke_candidates, 5)

```

# Lady Byng Memorial Trophy

This is for the player who has best sportsmanship while also having a high standard of playing. I will use low penalty minutes (PIM) as a proxy for that.

```{r}
lady_byng_candidates <- all_skaters %>%
  filter(GP >= 30) %>% 
  mutate(Sportsmanship = ifelse(PIM == 0, 1, 1 / PIM)) %>%
  arrange(desc(Sportsmanship)) %>%
  select(Player, Team, GP, PIM, Sportsmanship)

#ballot
lady_byng_top5 <- head(lady_byng_candidates, 5)

```

# Showing each ballot seprately

```{r}
## Hart Memorial Trophy
mvp_top5

## Vezina Trophy
vezina_top5

## Calder Memorial Trophy
calder_top5

## James Norris Memorial Trophy
norris_top5

## Frank J. Selke Trophy
selke_top5

## Lady Byng Memorial Trophy
lady_byng_top5
```

# The ballots in a single table

```{r}
ballots_combined <- bind_rows(
  mvp_top5 %>% mutate(Award = "Hart Memorial Trophy", Rank = row_number()),
  vezina_top5 %>% mutate(Award = "Vezina Trophy", Rank = row_number()),
  calder_top5 %>% mutate(Award = "Calder Memorial Trophy", Rank = row_number()),
  norris_top5 %>% mutate(Award = "James Norris Memorial Trophy", Rank = row_number()),
  selke_top5 %>% mutate(Award = "Frank J. Selke Trophy", Rank = row_number()),
  lady_byng_top5 %>% mutate(Award = "Lady Byng Memorial Trophy", Rank = row_number())
) %>%
  mutate(Voter = "Marwa") %>%
  select(Voter, Rank, Award, Player)

# Display the full table
ballots_combined %>% 
  kable("html") %>% 
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "hover", "condensed"))
```
